---
title: Electronic Resources Reporting - Set Up
author: Andy Meyer
date: '2017-09-07'
slug: electronic-resources-reporting-set-up
categories:
  - Electronic Resources
tags:
  - tidyverse
---

The setup file in my Electronic Resource report project is still a work in progress but it's the logical place to start this reflection and tour. There are two main sections: a place to load packages that I will need later on as well as a place to define the folders I will use for data import and data export.

## Loading Packages
Here is the R code for this section:
```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(mosaic)
library(zoo)
library(scales)
library(knitr)
```

The library function loads libraries that add new functions to base R. There are a lot of packages out there; I'm loading these because they are all required somewhere later in this project. 

## Setting up the Folders
```{r}
input.folder <- "C:/DataScience/inputs"
output.folder <- "C:/DataScience/outputs"
DB1folder <- "C:/DataScience/inputs/DB1Reports"
JR1folder <- "C:/DataScience/inputs/JR1Reports"
```

I'm still wondering if make specific definitions (like this) or setting up a working directory is the better way to specify where R should look for inputs and where it should look for outputs. I've defined these folders manually and that seems to make sense at the time; I might revisit this in the future and use a setwd() and make all of the folder pathways relative.